---
title: "Proteomic consistency metrics"
author: "Simina Boca, Shaojun Tang, Yi Bai, Nathan Edwards"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true # table of content true
    depth: 3  # specified by number of #
    ##css: vignette.css
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(knitr)
library(kableExtra)

options(warn=-1)
```

```{r, echo=FALSE, include=FALSE}
##Read in metrics for individual analytical samples and fractions and check number of fractions and samples

QCmetrics <- read.table("qcmetrics_PNNL.tsv", header=TRUE, sep="\t")
dim(QCmetrics)

head(QCmetrics)

##need to separate out the fractions if they are all equal to NA (this means they have not been pulled out before)
if(sum(!is.na(QCmetrics$fractionNum)) == 0)
{
  analyticalBasenameSplit <- strsplit(as.character(QCmetrics$analyticalBasename),"_Fr_")
  QCmetrics$analyticalBasename <- sapply(analyticalBasenameSplit, function(x){x[1]})
  QCmetrics$fractionNum <- as.numeric(sapply(analyticalBasenameSplit, function(x){x[2]}))
  unique(QCmetrics$fractionNum)
  QCmetrics$fractionNum[is.na(QCmetrics$fractionNum)] <- 25
}
```

```{r, echo=FALSE, include=FALSE}
##Get the fractions (check they are 1-24):

sort(unique(QCmetrics$fractionNum))
length(unique(QCmetrics$fractionNum))

QCmetrics$fractionNum[is.na(QCmetrics$fractionNum)] <- 25
```

```{r, echo=FALSE, include=FALSE}
##Get the number of unique analytical samples

analyticalSamples <- unique(QCmetrics$analyticalBasename)
length(analyticalSamples)

##Make sure the labels in spectrumBasename are unique

length(unique(QCmetrics$spectrumBasename))
```

```{r, echo=FALSE, include=FALSE}
##Change some column names to make things easier
QCmetrics <- rename(QCmetrics,
                    PrecursorIntensity.0 = minPrecursorIntensity,
                    PrecursorIntensity.5 = PrecursorIntensity5perc,
                    PrecursorIntensity.25 = PrecursorIntensity25perc,
                    PrecursorIntensity.50 = PrecursorIntensity50perc,
                    PrecursorIntensity.75 = PrecursorIntensity75perc,
                    PrecursorIntensity.95 = PrecursorIntensity95perc,
                    PrecursorIntensity.100 = maxPrecursorIntensity)
QCmetrics <- rename(QCmetrics,
                    PrecursorMZ.0 = minPrecursorMZ,
                    PrecursorMZ.5 = PrecursorMZ5perc,
                    PrecursorMZ.25 = PrecursorMZ25perc,
                    PrecursorMZ.50 = PrecursorMZ50perc,
                    PrecursorMZ.75 = PrecursorMZ75perc,
                    PrecursorMZ.95 = PrecursorMZ95perc,
                    PrecursorMZ.100 = maxPrecursorMZ)

##Create data frame melting the precursor intensities
QCmetricsLongPrecInt <- melt(QCmetrics[,c(2,3,grep("PrecursorIntensity.",colnames(QCmetrics)))],
                             variable.name="PrecursorIntensity",
                             value.name="value",
                             id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongPrecInt)
QCmetricsLongPrecInt <- cbind(QCmetricsLongPrecInt[,1:2],
                              colsplit(QCmetricsLongPrecInt[,3], "\\.",
                                       c("PrecursorIntensity","Percentile")),
                              QCmetricsLongPrecInt[,4])
head(QCmetricsLongPrecInt)
colnames(QCmetricsLongPrecInt)[5] <- "value"

##Create data frame melting the precursor MZ values
QCmetricsLongPrecMZ <- melt(QCmetrics[,c(2,3,grep("PrecursorMZ.",colnames(QCmetrics)))],
                            variable.name="PrecursorMZ",
                            value.name="value",
                            id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongPrecMZ)
QCmetricsLongPrecMZ <- cbind(QCmetricsLongPrecMZ[,1:2],
                             colsplit(QCmetricsLongPrecMZ[,3], "\\.",
                                      c("PrecursorMZ","Percentile")),
                             QCmetricsLongPrecMZ[,4])
head(QCmetricsLongPrecMZ)
colnames(QCmetricsLongPrecMZ)[5] <- "value"
```

# Number of MS2 spectra
### Per sample, across fraction (boxplot)
```{r, echo=FALSE, fig.height=6, fig.width=6}
##Get boxplot of number of spectra across fractions
g <- ggplot(QCmetrics, aes(y=numofMS2, x=analyticalBasename, col=analyticalBasename)) +
  geom_boxplot() +
  ggtitle("Number of MS2 spectra across fractions\nEach analytical sample is a different color") +
  xlab("Analytical sample") + ylab("Number of spectra") +
  scale_color_discrete(name="Analytical sample") +
  theme(axis.text.x  = element_blank())##element_text(angle=90, vjust=0.5)) 
if(length(analyticalSamples)<=10)
{
  g
} else {
  scale_colour_discrete(guide = FALSE)
}
```

### Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
g <- ggplot(QCmetrics, aes(y=numofMS2, x=fractionNum, col=analyticalBasename)) +
  geom_point() +
  stat_smooth(method="loess", fill="grey50", size=0)+##, color="grey50") +
  ##geom_line() +
  ggtitle("Number of MS2 spectra versus fraction\nEach analytical sample is a different color") +
  scale_color_discrete(name="Analytical sample") +
  xlab("Fraction") + ylab("Number of spectra") 
if(length(analyticalSamples)<=10)
{
  g
} else {
  scale_colour_discrete(guide = FALSE)
}
##scale_colour_discrete(guide = FALSE)
```

# Precursor intensity values
### Range of precursor intensity values
##### This table shows the range of the precursor intensity for each percentile, as a summary and to see if there are any values of 0 that cannot be log-transformed . 
```{r, echo = FALSE}
B = c(5,25,50,75,95)

rangetable = matrix(,nrow = 2, ncol = length(B), byrow = TRUE)
for (i in 1:length(B)){
  rangesub <-subset(QCmetricsLongPrecInt, QCmetricsLongPrecInt$Percentile == B[i])
  # filter to get only the wanted percentile
  #print(range(rangesub$value))
  rangetable[1,i] <- toString(round(range(rangesub$value),2))
  rangetable[2,i] <- toString(signif(log10(range(rangesub$value)),3))
}
colnames(rangetable)<-c("5th Perc","25th perc","50th perc","75th perc","95th perc")
rownames(rangetable)<-c("Original values", "log10 transformed values")
#rangetable
kable(rangetable) #%>% kable_styling(full_width = FALSE)

```

### Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.height=5, fig.width=12}
Pmin <- log10(min(QCmetricsLongPrecInt$value[QCmetricsLongPrecInt$value > 0]))
Pmax <- log10(max(QCmetricsLongPrecInt$value))

QCmetricsLongPrecInt$value2 <- QCmetricsLongPrecInt$value

QCmetricsLongPrecInt$value2[QCmetricsLongPrecInt$value2 == 0] <-
  min(QCmetricsLongPrecInt$value2[QCmetricsLongPrecInt$value2 > 0])*0.5

g <- ggplot(QCmetricsLongPrecInt[QCmetricsLongPrecInt$Percentile %in% c(5,25,50,75,95),], 
       aes(y=log10(value2), x=fractionNum, col=analyticalBasename, group=analyticalBasename)) +
  facet_grid(~ Percentile) +
  geom_point() +
  ylim(c(Pmin, Pmax)) +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("5%,25%,50%,75%,95% for precursor intensity\nEach analytical sample is a different color\nEach panel is a percentile") +
  scale_color_discrete(name="Analytical sample") +
  xlab("Fraction") + ylab("log10(Precursor intensity)") 
if(length(analyticalSamples)<=10)
{
  g
} else {
  scale_colour_discrete(guide = FALSE)
}
```

# Precursor m/z value 

### Per sample, versus fraction (scatterplot)

```{r, echo=FALSE, fig.height=5, fig.width=12}
g <- ggplot(QCmetricsLongPrecMZ[QCmetricsLongPrecMZ$Percentile %in% c(5,25,50,75,95),], 
       aes(y=value, x=fractionNum, col=analyticalBasename, group=analyticalBasename)) +
  facet_grid(~ Percentile) +
  geom_point() +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("5%,25%,50%,75%,95% for precursor m/z\nEach analytical sample is a different color\nEach panel is a percentile") +
  scale_color_discrete(name="Analytical sample") +
  xlab("Fraction") + ylab("Precursor m/z") 
if(length(analyticalSamples)<=10)
{
  g
} else {
  scale_colour_discrete(guide = FALSE)
}
```
