---
title: "Proteomic consistency metrics -- Development Version"
author: "Simina Boca, Shaojun Tang, Yi Bai, Nathan Edwards"
date: '`r format(Sys.Date(), "%B %d, %Y")`'       # adding current date 
output:
  html_document:                                  
    toc: true                                     # table of content : true
    toc_depth: 3                                  # level of content table
runtime: shiny
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggvis)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggvis)
library(RColorBrewer)
options(warn = -1)
```

```{r, echo=FALSE, include=FALSE}
##Read in metrics for individual analytical samples and fractions and check number of fractions and samples

QCmetrics <- read.table("Compref_BI_Proteome_QCmetrics.tsv", header=TRUE, sep="\t")
dim(QCmetrics)
head(QCmetrics)
```

```{r, echo=FALSE, include=FALSE}
##need to separate out the fractions if they are all equal to NA (this means they have not been pulled out before)
if(sum(!is.na(QCmetrics$fractionNum)) == 0)
{
  analyticalBasenameSplit <- strsplit(as.character(QCmetrics$analyticalBasename),"_Fr_")
  QCmetrics$analyticalBasename <- sapply(analyticalBasenameSplit, function(x){x[1]})
  QCmetrics$fractionNum <- as.numeric(sapply(analyticalBasenameSplit, function(x){x[2]}))
  unique(QCmetrics$fractionNum)
  QCmetrics$fractionNum[is.na(QCmetrics$fractionNum)] <- 25
}

##Get the fractions (check they are 1-24):

sort(unique(QCmetrics$fractionNum))
length(unique(QCmetrics$fractionNum))

QCmetrics$fractionNum[is.na(QCmetrics$fractionNum)] <- 25
```

```{r, echo=FALSE, include=FALSE}
##Get the number of unique analytical samples

analyticalSamples <- unique(QCmetrics$analyticalBasename)
length(analyticalSamples)

##Make sure the labels in spectrumBasename are unique

length(unique(QCmetrics$spectrumBasename))
```

```{r, echo=FALSE, include=FALSE}
##Change some column names to make things easier
QCmetrics <- rename(QCmetrics,
                    PrecursorIntensity.0 = minPrecursorIntensity,
                    PrecursorIntensity.5 = PrecursorIntensity5perc,
                    PrecursorIntensity.25 = PrecursorIntensity25perc,
                    PrecursorIntensity.50 = PrecursorIntensity50perc,
                    PrecursorIntensity.75 = PrecursorIntensity75perc,
                    PrecursorIntensity.95 = PrecursorIntensity95perc,
                    PrecursorIntensity.100 = maxPrecursorIntensity)
QCmetrics <- rename(QCmetrics,
                    PrecursorMZ.0 = minPrecursorMZ,
                    PrecursorMZ.5 = PrecursorMZ5perc,
                    PrecursorMZ.25 = PrecursorMZ25perc,
                    PrecursorMZ.50 = PrecursorMZ50perc,
                    PrecursorMZ.75 = PrecursorMZ75perc,
                    PrecursorMZ.95 = PrecursorMZ95perc,
                    PrecursorMZ.100 = maxPrecursorMZ)

##Create data frame melting the precursor intensities
QCmetricsLongPrecInt <- melt(QCmetrics[,c(2,3,grep("PrecursorIntensity.",colnames(QCmetrics)))],
                             variable.name="PrecursorIntensity",
                             value.name="value",
                             id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongPrecInt)
QCmetricsLongPrecInt <- cbind(QCmetricsLongPrecInt[,1:2],
                              colsplit(QCmetricsLongPrecInt[,3], "\\.",
                                       c("PrecursorIntensity","Percentile")),
                              QCmetricsLongPrecInt[,4])
head(QCmetricsLongPrecInt)
colnames(QCmetricsLongPrecInt)[5] <- "value"

##Create data frame melting the precursor MZ values
QCmetricsLongPrecMZ <- melt(QCmetrics[,c(2,3,grep("PrecursorMZ.",colnames(QCmetrics)))],
                            variable.name="PrecursorMZ",
                            value.name="value",
                            id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongPrecMZ)
QCmetricsLongPrecMZ <- cbind(QCmetricsLongPrecMZ[,1:2],
                             colsplit(QCmetricsLongPrecMZ[,3], "\\.",
                                      c("PrecursorMZ","Percentile")),
                             QCmetricsLongPrecMZ[,4])
head(QCmetricsLongPrecMZ)
colnames(QCmetricsLongPrecMZ)[5] <- "value"
```

```{r, echo = FALSE}
# generating a color Palettes for ggvis 
plotcolor <- brewer.pal(length(analyticalSamples), "Dark2")
```


## Number of MS2 spectra
### Per sample, across fraction (boxplot)

```{r, echo=FALSE, fig.height=8, fig.width=6}
##Get boxplot of number of spectra across fractions
g <- ggplot(QCmetrics, aes(y=numofMS2, x=analyticalBasename, col=analyticalBasename)) +
  geom_boxplot(colour = plotcolor[1:length(analyticalSamples)]) +
  ggtitle("Number of MS2 spectra across fractions\nEach analytical sample is a different color") +
  xlab("Analytical sample") + ylab("Number of spectra") +
  scale_color_discrete(name = "Analytical sample") +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
if(length(analyticalSamples)<=10)
{
  g
} else {
  scale_colour_discrete(guide = FALSE)
}

```


<a href="#top">Back to top</a>

### Per sample, across fraction (interactive boxplot)
```{r, echo = FALSE, fig.height=8, fig.width=7}

#adding a keys column to the QC table to access the key
    QC1 <- QCmetrics
    QC1$id <- 1:nrow(QC1)
    
    showSamps <- function(x){if(is.null(x)) return(NULL);
     row <- QC1[QC1$id == x$id,]
     paste0("Analytical sample:", format(row[2])," ", "Fraction:",format(row[3]))}
    
   # analytical_labels <- factor(c(unique(QC1$analyticalBasename)), labels = c(as.character(analyticalSamples)))
    

  QC1 %>%
      ggvis(x= ~as.numeric(analyticalBasename), y= ~numofMS2, opacity:=0.5,fill=~analyticalBasename) %>%
      #layer_model_predictions(model='loess') %>%
      add_tooltip(showSamps, "click") %>%
      layer_boxplots() %>%
      layer_points(key := ~ id) %>%
      add_axis("x", title = "Analytical sample", title_offset = 77, values =  seq(as.character(analyticalSamples)), 
           properties = axis_props(labels=list(angle=90, align = "left", baseline = "middle"))) %>%
      #add_axis("x", title = "Analytical sample",title_offset = 77,
           #    properties = axis_props(labels=list(angle=90,align = "left", baseline = "middle"))) %>% 
      add_axis("y", title = "Number of spectra", title_offset = 90) %>% 
      hide_legend("fill") %>%
     # set_options(height = 680, width = 550)
      scale_nominal("fill", range = plotcolor[1:length(analyticalSamples)])
```


<a href="#top">Back to top</a>


### Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
g <- ggplot(QCmetrics, aes(y=numofMS2, x=fractionNum, col=analyticalBasename)) +
  geom_point() +
  scale_color_brewer(palette ="Dark2", name = "Analytical sample") +
  stat_smooth(method="loess", fill="grey50", size=0)+##, color="grey50") +
  ##geom_line() +
  ggtitle("Number of MS2 spectra versus fraction\nEach analytical sample is a different color") +
  xlab("Fraction") + ylab("Number of spectra") ##+
  ##scale_colour_discrete(name = "Analytical sample")
if(length(analyticalSamples)<=10)
{
  g
} else {
  scale_colour_discrete(guide = FALSE)
}
```



<a href="#top">Back to top</a>


Number of MS2 spectra versus fraction (boxplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
ggplot(QCmetrics, aes(y=numofMS2, x=factor(fractionNum))) +
  geom_point(aes(col=analyticalBasename)) +
  stat_smooth(aes(y=numofMS2, x=fractionNum, col=analyticalBasename),method="loess", fill="grey50", size=0)+##, color="grey50") +
  geom_boxplot(outlier.size=NA) +
  scale_color_brewer(palette ="Dark2", name="Analytical sample") +
  ggtitle("Number of MS2 spectra versus fraction\nEach analytical sample is a different color") +
  xlab("Fraction") + ylab("Number of spectra") 
##scale_colour_discrete(guide = FALSE)
```


<a href="#top">Back to top</a>



## Precursor Intensity values
### Range of Precursor intensity values
This table shows the range of the precursor intensity for each percentile, as a summary and to see if there are any value of 0 that cannot be log-trnasformed.

```{r, echo = FALSE}
B = c(5,25,50,75,95)

rangetable = matrix(,nrow = 2, ncol = length(B), byrow = TRUE)
for (i in 1:length(B)){
  rangesub <-subset(QCmetricsLongPrecInt, QCmetricsLongPrecInt$Percentile == B[i])
  # filter to get only the wanted percentile
  #print(range(rangesub$value))
  rangetable[1,i] <- toString(round(range(rangesub$value),2))
  rangetable[2,i] <- toString(signif(log10(range(rangesub$value)),3))
}
colnames(rangetable)<-c("5th Perc","25th perc","50th perc","75th perc","95th perc")
rownames(rangetable)<-c("range", "range of log10 transformed")
#rangetable
kable(rangetable) #%>% kable_styling(full_width = FALSE)

```


<a href="#top">Back to top</a>


###Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
Pmin <- log10(min(QCmetricsLongPrecInt$value[QCmetricsLongPrecInt$value > 0]))
Pmax <- log10(max(QCmetricsLongPrecInt$value))

QCmetricsLongPrecInt$value2 <- QCmetricsLongPrecInt$value
QCmetricsLongPrecInt$value2[QCmetricsLongPrecInt$value2 == 0] <-
  min(QCmetricsLongPrecInt$value2[QCmetricsLongPrecInt$value2 > 0])*0.5

g <- ggplot(QCmetricsLongPrecInt[QCmetricsLongPrecInt$Percentile %in% c(5,25,50,75,95),], 
       aes(y=log10(value2), x=fractionNum, col=analyticalBasename, group=analyticalBasename)) + 
  facet_grid(~ Percentile) +
  geom_point() +
  ylim(c(Pmin, Pmax)) +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("5%,25%,50%,75%,95% for precursor intensity\nEach analytical sample is a different color\nEach panel is a percentile") +
  xlab("Fraction") + ylab("log10(Precursor intensity)") + 
  ##scale_colour_discrete(name = "Analytical sample") +
 scale_color_brewer(palette ="Dark2", name="Analytical sample")
if(length(analyticalSamples)<=10)
{
  g
} else {
  scale_colour_discrete(guide = FALSE)
}
```

<a href="#top">Back to top</a>

### Per sample, versus fraction (interactive scatterplot)
```{r, echo = FALSE}
shinyApp(
  ui = fluidPage(
    sidebarLayout(
    sidebarPanel(
      selectInput("percentile", "Percentile:", 
                  choices=c(5,25,50,75,95),
                  selected = 50),
      tags$hr()
    ),
    mainPanel(
      ggvisOutput("PrecursorInt"),
      uiOutput("PrecursorInt_ui")
    )
  )
  ),
  
  server = function(input, output){
    
  QCmetricsLong <- reactive({
    QC4 <- as.data.frame(QCmetricsLongPrecInt)
    QC2 <- QCmetricsLongPrecInt[QCmetricsLongPrecInt$Percentile %in% c(5,25,50,75,95),]
    
    QC4 <- QC2[QC2$Percentile %in% c(input$percentile),]
   
    #QC4$keys = 1:nrow(QC4)
    
    #QC4
  })
  
  observe({
    
    QC4 <- as.data.frame(QCmetricsLong())  
    QC4$keys = 1:nrow(QC4)

    
    showSamps <- function(x){if(is.null(x)) return(NULL);
      paste0(format(QC4[QC4$keys == x$keys,
                        c("analyticalBasename")]))}
    
    QC4 %>%
      ggvis(~fractionNum, ~log10(value2),opacity:=0.5,fill=~analyticalBasename) %>%
      layer_model_predictions(model='loess') %>%
      add_tooltip(html=showSamps, on="click") %>%
      layer_points(key := ~ keys) %>%
      add_axis("x", title = "Fraction") %>% 
      add_axis("y", title = "log10(Precursor intensity)") %>% 
      hide_legend("fill") %>%
      scale_numeric("y", domain = c(Pmin,Pmax)) %>%
      scale_nominal("fill", range = plotcolor[1:length(analyticalSamples)]) %>%
      bind_shiny("PrecursorInt","PrecursorInt_ui")
  })
    
  }
  
)
```


<a href="#top">Back to top</a>


### Percentage of 0 precursor intensities at different percentiles
In this table, each row represents a fraction and each column represents a percentile. Each value represents the percent of samples that have a precursor intensity of 0 at that fraction and that percentile.
```{r,echo = FALSE}
# This table indicates the percente of 0 in the value column

#generate a dataframe contians only the precInt is 0
QCPrecIntof0 <- subset(QCmetricsLongPrecInt, QCmetricsLongPrecInt$value == 0)
#QCPrecIntof0

#generate the dataframe where the rows are different fractions and the columns are different percentile
A=c(1:25)
B=c(5,25,50,75,95)
QCPrecIntTable=matrix(,nrow = length(A), ncol = length(B))
for (i in 1:length(A)){
  for (j in 1:length(B)){
    QCPrecIntTable[i,j]<-round(length(which(QCPrecIntof0$Percentile == B[j] & QCPrecIntof0$fractionNum == i))/length(which(QCmetricsLongPrecInt$fractionNum == i & QCmetricsLongPrecInt$Percentile == B[j]))*100, 3)
  }
}
#QCPrecIntTable
colnames(QCPrecIntTable) <- c("5th Perc", "25th Perc", "50th Perc", "75th Perc","95th Perc")
rownames(QCPrecIntTable) <- c(paste0("fraction", c(1:25)))
kable(QCPrecIntTable) #%>% kable_styling(bootstrap_options = "striped", full_width = F)

```


<a href="#top">Back to top</a>



## Precursor m/z value
### Per sample, versus fraction (scatterplot)

```{r, echo=FALSE, fig.width=10, fig.height=6}
g <- ggplot(QCmetricsLongPrecMZ[QCmetricsLongPrecMZ$Percentile %in% c(5,25,50,75,95),], 
       aes(y=value, x=fractionNum, col=analyticalBasename, group=analyticalBasename)) +
  facet_grid(~ Percentile) +
  geom_point() +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("5%,25%,50%,75%,95% for precursor m/z\nEach analytical sample is a different color\nEach panel is a percentile") +
  ##scale_color_discrete(name="Analytical sample") +
  scale_color_brewer(palette ="Dark2", name="Analytical sample") +
  xlab("Fraction") + ylab("Precursor m/z")
if(length(analyticalSamples)<=10)
{
  g
} else {
  scale_colour_discrete(guide = FALSE)
}
```

<a href="#top">Back to top</a>


### Per sample, versus fraction (interactive scatterplot)
```{r, echo = FALSE, fig.width=10, fig.height=10}
shinyApp(
  ui = fluidPage(
    sidebarLayout(
    sidebarPanel(
      selectInput("percentile", "Percentile:", 
                  choices=c(5,25,50,75,95),
                  selected = 50),
      tags$hr()
    ),
    mainPanel(
      ggvisOutput("PrecursorMZ"),
      uiOutput("PrecursorMZ_ui")
    )
  )
  ),
  
  server = function(input, output){
    
  QCmetricsLong <- reactive({
    QC3 <- as.data.frame(QCmetricsLongPrecMZ)
    QC2 <- QCmetricsLongPrecMZ [QCmetricsLongPrecMZ $Percentile %in% c(5,25,50,75,95),]
    
    QC3 <- QC3[QC3$Percentile %in% c(input$percentile),]
   
    #QC4$keys = 1:nrow(QC4)
    
    #QC4
  })
  
  observe({
    
    QC3 <- as.data.frame(QCmetricsLong())  
    QC3$keys = 1:nrow(QC3)

    
    showSamps <- function(x){if(is.null(x)) return(NULL);
      paste0(format(QC3[QC3$keys == x$keys,
                        c("analyticalBasename")]))}
    
    QC3 %>%
      ggvis(~fractionNum, ~value,opacity:=0.5,fill=~analyticalBasename) %>%
      layer_model_predictions(model='loess') %>%
      add_tooltip(html=showSamps, on="click") %>%
      layer_points(key := ~ keys) %>%
      add_axis("x", title = "Fraction") %>% 
      add_axis("y", title = "Precursor M/Z", title_offset = 50) %>% 
      hide_legend("fill") %>%
      #scale_numeric("y", domain = c(min(QCmetricsLongPrecMZ$value)-10,max(QCmetricsLongPrecMZ$value))) %>%   
      scale_nominal("fill", range = plotcolor[1:length(analyticalSamples)]) %>%
      bind_shiny("PrecursorMZ","PrecursorMZ_ui")
  })
    
  }
  
)
```
<a href="#top">Back to top</a>