---
title: "Proteomic consistency metrics -- Development Version with static metrics"
author: "Simina Boca, Shaojun Tang, Yi Bai, Nathan Edwards"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: 3
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggvis)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggvis)
library(RColorBrewer)
library(pander)      ## this package is used ot convert the .md file into PDF
options(warn = -1)
```

```{r, echo=FALSE, include=FALSE}
##Read in metrics for individual analytical samples and fractions and check number of fractions and samples

QCmetrics <- read.table("Compref_BI_Proteome_QCmetrics.tsv", header=TRUE, sep="\t")
dim(QCmetrics)
head(QCmetrics)
```

```{r, echo=FALSE, include=FALSE}
##need to separate out the fractions if they are all equal to NA (this means they have not been pulled out before)
if(sum(!is.na(QCmetrics$fractionNum)) == 0)
{
  analyticalBasenameSplit <- strsplit(as.character(QCmetrics$analyticalBasename),"_Fr_")
  QCmetrics$analyticalBasename <- sapply(analyticalBasenameSplit, function(x){x[1]})
  QCmetrics$fractionNum <- as.numeric(sapply(analyticalBasenameSplit, function(x){x[2]}))
  unique(QCmetrics$fractionNum)
  QCmetrics$fractionNum[is.na(QCmetrics$fractionNum)] <- 25
}

##Get the fractions (check they are 1-24):

sort(unique(QCmetrics$fractionNum))
length(unique(QCmetrics$fractionNum))

QCmetrics$fractionNum[is.na(QCmetrics$fractionNum)] <- 25
```

```{r, echo=FALSE, include=FALSE}
##Get the number of unique analytical samples

analyticalSamples <- unique(QCmetrics$analyticalBasename)
length(analyticalSamples)

##Make sure the labels in spectrumBasename are unique

length(unique(QCmetrics$spectrumBasename))
```

```{r, echo=FALSE, include=FALSE}
##Change some column names to make things easier
QCmetrics <- rename(QCmetrics,
                    PrecursorIntensity.0 = minPrecursorIntensity,
                    PrecursorIntensity.5 = PrecursorIntensity5perc,
                    PrecursorIntensity.25 = PrecursorIntensity25perc,
                    PrecursorIntensity.50 = PrecursorIntensity50perc,
                    PrecursorIntensity.75 = PrecursorIntensity75perc,
                    PrecursorIntensity.95 = PrecursorIntensity95perc,
                    PrecursorIntensity.100 = maxPrecursorIntensity)
QCmetrics <- rename(QCmetrics,
                    PrecursorMZ.0 = minPrecursorMZ,
                    PrecursorMZ.5 = PrecursorMZ5perc,
                    PrecursorMZ.25 = PrecursorMZ25perc,
                    PrecursorMZ.50 = PrecursorMZ50perc,
                    PrecursorMZ.75 = PrecursorMZ75perc,
                    PrecursorMZ.95 = PrecursorMZ95perc,
                    PrecursorMZ.100 = maxPrecursorMZ)

QCmetrics <- rename(QCmetrics,
                    PrecursorMW.0 = minPrecursorMW,
                    PrecursorMW.5 = PrecursorMW5perc,
                    PrecursorMW.25 = PrecursorMW25perc,
                    PrecursorMW.50 = PrecursorMW50perc,
                    PrecursorMW.75 = PrecursorMW75perc,
                    PrecursorMW.95 = PrecursorMW95perc,
                    PrecursorMW.100 = maxPrecursorMW)

QCmetrics <- rename(QCmetrics,
                    MS2perMS.0 = minMS2perMS,
                    MS2perMS.5 = MS2perMS5perc,
                    MS2perMS.25 = MS2perMS25perc,
                    MS2perMS.50 = MS2perMS50perc,
                    MS2perMS.75 = MS2perMS75perc,
                    MS2perMS.95 = MS2perMS95perc,
                    MS2perMS.100 = maxMS2perMS)

QCmetrics <- rename(QCmetrics,
                    MS2Charge.1.toCharge2Ratio = MS2Charge1toCharge2Ratio,
                    MS2Charge.3.toCharge2Ratio = MS2Charge3toCharge2Ratio,
                    MS2Charge.4.toCharge2Ratio = MS2Charge4toCharge2Ratio)


##Create data frame melting the precursor intensities
QCmetricsLongPrecInt <- melt(QCmetrics[,c(2,3,grep("PrecursorIntensity.",colnames(QCmetrics)))],
                             variable.name="PrecursorIntensity",
                             value.name="value",
                             id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongPrecInt)
QCmetricsLongPrecInt <- cbind(QCmetricsLongPrecInt[,1:2],
                              colsplit(QCmetricsLongPrecInt[,3], "\\.",
                                       c("PrecursorIntensity","Percentile")),
                              QCmetricsLongPrecInt[,4])
head(QCmetricsLongPrecInt)
colnames(QCmetricsLongPrecInt)[5] <- "value"

##Create data frame melting the precursor MZ values
QCmetricsLongPrecMZ <- melt(QCmetrics[,c(2,3,grep("PrecursorMZ.",colnames(QCmetrics)))],
                            variable.name="PrecursorMZ",
                            value.name="value",
                            id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongPrecMZ)
QCmetricsLongPrecMZ <- cbind(QCmetricsLongPrecMZ[,1:2],
                             colsplit(QCmetricsLongPrecMZ[,3], "\\.",
                                      c("PrecursorMZ","Percentile")),
                             QCmetricsLongPrecMZ[,4])
head(QCmetricsLongPrecMZ)
colnames(QCmetricsLongPrecMZ)[5] <- "value"

##Create data frame melting the precursor MW values
QCmetricsLongPrecMW <- melt(QCmetrics[,c(2,3,grep("PrecursorMW.",colnames(QCmetrics)))],
                            variable.name="PrecursorMW",
                            value.name="value",
                            id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongPrecMW)
QCmetricsLongPrecMW <- cbind(QCmetricsLongPrecMW[,1:2],
                             colsplit(QCmetricsLongPrecMW[,3], "\\.",
                                      c("PrecursorMW","Percentile")),
                             QCmetricsLongPrecMW[,4])
head(QCmetricsLongPrecMW)
colnames(QCmetricsLongPrecMW)[5] <- "value"


##Create data frame melting the MS2perMS values
QCmetricsLongMS2perMS <- melt(QCmetrics[,c(2,3,grep("MS2perMS.",colnames(QCmetrics)))],
                              variable.name=" MS2perMS",
                              value.name="value",
                              id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongMS2perMS)
QCmetricsLongMS2perMS <- cbind(QCmetricsLongMS2perMS[,1:2],
                               colsplit(QCmetricsLongMS2perMS[,3], "\\.",
                                        c("MS2perMS","Percentile")),
                               QCmetricsLongMS2perMS[,4])
head(QCmetricsLongMS2perMS)
colnames(QCmetricsLongMS2perMS)[5] <- "value"

##Create data frame melting MS2 charge ratio
QCmetricsLongMS2ChargeRatio <- melt(QCmetrics[,c(2,3,grep("MS2Charge.",colnames(QCmetrics)))],
                                    variable.name=" MS2ChargeRatio",
                                    value.name="value",
                                    id.vars=c("analyticalBasename","fractionNum"))
head(QCmetricsLongMS2ChargeRatio)
QCmetricsLongMS2ChargeRatio <- cbind(QCmetricsLongMS2ChargeRatio[,1:2],
                                     colsplit(QCmetricsLongMS2ChargeRatio[,3], "\\.",
                                              c("MS2charge","ChargeState","toCharge2ratio")),
                                     QCmetricsLongMS2ChargeRatio[,4])
head(QCmetricsLongMS2ChargeRatio)
colnames(QCmetricsLongMS2ChargeRatio)[6] <- "value"

```

```{r, echo = FALSE}
# generating a color Palettes for ggvis 
# plotColorPalettes is the function for generating the pallettes, the colors in the palettes depends on the number we put in 
# in our case, the palettes is called plotcolor

plotColorPalettes <- function(g){
  d <- 360/g
  h <- cumsum(c(15, rep(d,g - 1)))
  hcl(h = h, c = 100, l = 65)
}
plotcolor <- plotColorPalettes(length(analyticalSamples)) # the number of colors is the same as how many analytical smples in the data. 
```


## Number of MS2 spectra
### Per sample, across fraction (boxplot)

```{r, echo=FALSE, fig.height=8, fig.width=6}
##Get boxplot of number of spectra across fractions
g <- ggplot(QCmetrics, aes(y=numofMS2, x=analyticalBasename, col=analyticalBasename)) +
  geom_boxplot(colour = plotcolor[1:length(analyticalSamples)]) +
  ggtitle("Number of MS2 spectra across fractions\nEach analytical sample is a different color") +
  xlab("Analytical sample") + ylab("Number of spectra") +
  scale_color_discrete(name = "Analytical sample") +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}

```

<a href="#top">Back to top</a>
  
  
### Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
g <- ggplot(QCmetrics, aes(y=numofMS2, x=fractionNum, col=analyticalBasename)) +
  geom_point() +
  scale_color_manual (values = c(plotcolor), name = "Analytical Sample") +
  stat_smooth(method="loess", fill="grey50", size=0)+##, color="grey50") +
  ##geom_line() +
  ggtitle("Number of MS2 spectra versus fraction\nEach analytical sample is a different color") +
  xlab("Fraction") + ylab("Number of spectra") ##+
##scale_colour_discrete(name = "Analytical sample")
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}
```

<a href="#top">Back to top</a>
  
  
### Number of MS2 spectra versus fraction (boxplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
ggplot(QCmetrics, aes(y=numofMS2, x=factor(fractionNum))) +
  geom_point(aes(col=analyticalBasename)) +
  stat_smooth(aes(y=numofMS2, x=fractionNum, col=analyticalBasename),method="loess", fill="grey50", size=0)+##, color="grey50") +
  geom_boxplot(outlier.size=NA) +
  scale_color_manual (values = c(plotcolor), name = "Analytical Sample") +
  ggtitle("Number of MS2 spectra versus fraction\nEach analytical sample is a different color") +
  xlab("Fraction") + ylab("Number of spectra") 
##g + theme(legend.position="none")
```

<a href="#top">Back to top</a>
  
  
## Precursor intensity values
### Range of Precursor intensity values
This table shows the range of the precursor intensity for each percentile, as a summary and to see if there are any value of 0 that cannot be log-trnasformed.

```{r, echo = FALSE}
B = c(5,25,50,75,95)

rangetable = matrix(,nrow = 2, ncol = length(B), byrow = TRUE)
for (i in 1:length(B)){
  rangesub <-subset(QCmetricsLongPrecInt, QCmetricsLongPrecInt$Percentile == B[i])
  # filter to get only the wanted percentile
  #print(range(rangesub$value))
  rangetable[1,i] <- toString(round(range(rangesub$value),2))
  rangetable[2,i] <- toString(signif(log10(range(rangesub$value)),3))
}
colnames(rangetable)<-c("5th Perc","25th perc","50th perc","75th perc","95th perc")
rownames(rangetable)<-c("range", "range of log10 transformed")
#rangetable
kable(rangetable) #%>% kable_styling(full_width = FALSE)

```

<a href="#top">Back to top</a>
  
  
###Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
Pmin <- log10(min(QCmetricsLongPrecInt$value[QCmetricsLongPrecInt$value > 0]))
Pmax <- log10(max(QCmetricsLongPrecInt$value))

QCmetricsLongPrecInt$value2 <- QCmetricsLongPrecInt$value
QCmetricsLongPrecInt$value2[QCmetricsLongPrecInt$value2 == 0] <-
  min(QCmetricsLongPrecInt$value2[QCmetricsLongPrecInt$value2 > 0])*0.5

g <- ggplot(QCmetricsLongPrecInt[QCmetricsLongPrecInt$Percentile %in% c(5,25,50,75,95),], 
            aes(y=log10(value2), x=fractionNum, col=analyticalBasename, group=analyticalBasename)) + 
  facet_grid(~ Percentile) +
  geom_point() +
  ylim(c(Pmin, Pmax)) +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("5%,25%,50%,75%,95% for precursor intensity\nEach analytical sample is a different color\nEach panel is a percentile") +
  xlab("Fraction") + ylab("log10(Precursor intensity)") + 
  ##scale_colour_discrete(name = "Analytical sample") +
  scale_color_manual (values = c(plotcolor), name = "Analytical Sample")
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}
```

<a href="#top">Back to top</a>
  
  
### Percentage of 0 precursor intensities at different percentiles
In this table, each row represents a fraction and each column represents a percentile. Each value represents the percent of samples that have a precursor intensity of 0 at that fraction and that percentile.
```{r,echo = FALSE}
# This table indicates the percente of 0 in the value column

#generate a dataframe contians only the precInt is 0
QCPrecIntof0 <- subset(QCmetricsLongPrecInt, QCmetricsLongPrecInt$value == 0)
#QCPrecIntof0

#generate the dataframe where the rows are different fractions and the columns are different percentile
A=c(1:25)
B=c(5,25,50,75,95)
QCPrecIntTable=matrix(,nrow = length(A), ncol = length(B))
for (i in 1:length(A)){
  for (j in 1:length(B)){
    QCPrecIntTable[i,j]<-round(length(which(QCPrecIntof0$Percentile == B[j] & QCPrecIntof0$fractionNum == i))/length(which(QCmetricsLongPrecInt$fractionNum == i & QCmetricsLongPrecInt$Percentile == B[j]))*100, 3)
  }
}
#QCPrecIntTable
colnames(QCPrecIntTable) <- c("5th Perc", "25th Perc", "50th Perc", "75th Perc","95th Perc")
rownames(QCPrecIntTable) <- c(paste0("fraction", c(1:25)))
kable(QCPrecIntTable) #%>% kable_styling(bootstrap_options = "striped", full_width = F)

```

<a href="#top">Back to top</a>
  
  
## Precursor m/z values
### Per sample, versus fraction (scatterplot)
  
```{r, echo=FALSE, fig.width=10, fig.height=6}
g <- ggplot(QCmetricsLongPrecMZ[QCmetricsLongPrecMZ$Percentile %in% c(5,25,50,75,95),], 
            aes(y=value, x=fractionNum, col=analyticalBasename, group=analyticalBasename)) +
  facet_grid(~ Percentile) +
  geom_point() +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("5%,25%,50%,75%,95% for precursor m/z\nEach analytical sample is a different color\nEach panel is a percentile") +
  ##scale_color_discrete(name="Analytical sample") +
  scale_color_manual (values = c(plotcolor), name = "Analytical Sample") +
  xlab("Fraction") + ylab("Precursor m/z")
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}
```

<a href="#top">Back to top</a>
  
  
## Precursor Molecular weight (MW) values
### Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
g <- ggplot(QCmetricsLongPrecMW[QCmetricsLongPrecMW$Percentile %in% c(5,25,50,75,95),], 
            aes(y=value, x=fractionNum, col=analyticalBasename, group=analyticalBasename)) +
  facet_grid(~ Percentile) +
  geom_point() +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("5%,25%,50%,75%,95% for precursor molecular weight\nEach analytical sample is a different color\nEach panel is a percentile") +
  ##scale_color_discrete(name="Analytical sample") +
  scale_color_manual (values = c(plotcolor), name = "Analytical Sample") +
  xlab("Fraction") + ylab("Precursor Molecular Weight")
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}
```

<a href="#top">Back to top</a>
  
  
### Per sample, across fraction at 50 percentile (boxplot)
```{r, echo=FALSE, fig.height=8, fig.width=6}
##Get boxplot of number of spectra across fractions
g <- ggplot(QCmetricsLongPrecMW[QCmetricsLongPrecMW$Percentile %in% c(50),], aes(y=value, x=analyticalBasename, col=analyticalBasename)) +
  geom_boxplot(colour = plotcolor[1:length(analyticalSamples)]) +
  ggtitle("Molecular weight at the 50 percentile across fractions\nEach analytical sample is a different color") +
  xlab("Analytical sample") + ylab("Precursor Molecular Weight") +
  scale_color_discrete(name = "Analytical sample") +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}

```

<a href="#top">Back to top</a>
  
  
### Per sample, across fraction (boxplot)
```{r, echo=FALSE, fig.width=10, fig.height= 8}
g <- ggplot(QCmetricsLongPrecMW[QCmetricsLongPrecMW$Percentile %in% c(5,25,50,75,95),], 
            aes(y=value, x=analyticalBasename, col=analyticalBasename, group=analyticalBasename)) +
  facet_grid(~ Percentile) +
  geom_boxplot () +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("5%,25%,50%,75%,95% for precursor molecular weight\nEach analytical sample is a different color\nEach panel is a percentile") +
  ##scale_color_discrete(name="Analytical sample") +
  scale_color_manual (values = c(plotcolor), name = "Analytical Sample") +
  xlab("Fraction") + ylab("Precursor Molecular Weight") +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5)) +
  theme(legend.position="top")
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}
```

<a href="#top">Back to top</a>


## MS2 per MS values
### Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
g <- ggplot(QCmetricsLongMS2perMS[QCmetricsLongMS2perMS$Percentile %in% c(5,25,50,75,95),], 
            aes(y=value, x=fractionNum, col=analyticalBasename, group=analyticalBasename)) +
  facet_grid(~ Percentile) +
  geom_point() +
  ##geom_line() +
  geom_smooth(method = loess, span = 0.3) +
  ggtitle("5%,25%,50%,75%,95% for MS2 per MS values\nEach analytical sample is a different color\nEach panel is a percentile") +
  ##scale_color_discrete(name="Analytical sample") + 
  scale_color_manual (values = c(plotcolor), name = "Analytical Sample") +
  xlab("Fraction") + ylab("MS2perMS value") +
  theme(legend.position="top")
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}
```

<a href="#top">Back to top</a>

  
## Different MS2 charge state to MS2 charge 2 ratio
### Per sample, versus fraction (scatterplot)
```{r, echo=FALSE, fig.width=10, fig.height=6}
g <- ggplot(QCmetricsLongMS2ChargeRatio[QCmetricsLongMS2ChargeRatio$ChargeState %in% c(1,3,4),], 
            aes(y=value, x=fractionNum, col=analyticalBasename, group=analyticalBasename)) +
  facet_grid(~ ChargeState) +
  geom_point() +
  ##geom_line() +
  ##geom_smooth() +
  ggtitle("MS2 charge1,3,4 to MS2 charge2 ratio\nEach analytical sample is a different color\nEach panel is a different MS2 charge state") +
  ##scale_color_discrete(name="Analytical sample") + 
  scale_color_manual (values = c(plotcolor), name = "Analytical Sample") +
  xlab("Fraction") + ylab("Charge Ratio") +
  ylim(0,max(QCmetricsLongMS2ChargeRatio$value)+0.5)
if(length(analyticalSamples)<=10)
{
  g
} else {
  g + theme(legend.position="none")
}
```

<a href="#top">Back to top</a>


## Data file 
We used the "Compref_BI_Proteome_QCmetrics.tsv" data file in this report.
To see the data file in tsv format, please click 
<a href="./Compref_BI_Proteome_QCmetrics.tsv" target="_blank"><b>Here</b></a>


<a href="#top">Back to top</a>


## PDF file 
To generate the PDF file of this report please run the following code in the R command line 
"rmarkdown::render("./developed_version_static_7:5.Rmd", output_format="pdf_document")"
To view the generated PDF file, please click
<a href="./developed_version_static_7_5.pdf" target="_blank"><b>Here</b></a>


<a href="#top">Back to top</a>

